Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-15 2017-Xudong-SMT/FM-2018-extended.tex
new OT ex

\item[$\bullet$] $S_j \subseteq \AlgA$ is a set of action terms,
  denoting the $\Sort$\footnote{The formal definition of $Sorts$ (set of actions of a Hole
  or pNet), $Leaves$ and $Holes$ (all pLTSs (resp holes) in a pNet
  hierarchical system. These can be found in \cite{henrio:Forte2016}.}
  of hole $j$.

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions
of several LTSs with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:lotos-composed}
%
%The \texttt{LOTOS} pNet of Fig. \ref{schema:lotos-pnet} has 2
%controllers and 2 holes. We will show its full open Automaton in
%section \ref{section:full-result}. We give one of its transitions here:
%
% \smallskip
% $ ot_2  = \openrule{
%                       \{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~
%                       \{\xrightarrow{hb_{12}}_P\}, ~~ 
%                        [s_0=0 \!\land\! hb_{12}=\delta(x)], ~~ 
%                       \{s_0 \xleftarrow{} 1\}     
%                      }
%     {\ostate{00} \xrightarrow{\underline{\delta(x)}} \ostate{01}}$
\end{example}
%
%The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.
%
%The transition in the conclusion of this rule starts in
%state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
%states. It ends in a different structured state, because in $ot_2$
%both pLTS conduct transitions, showing that the control has gone from
%processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
%$acc(x)$. Controller $C_1$ has only one state, but its state variable
%$s_0$ has changed to 1.

\paragraph{Structural Semantic Rules:}
\TODO{Do we push the rules in the appendix, and leave only a short high
  level desciption here ?}

Now we build the semantics of an open pNet as an open automaton where
LTSs are the pLTSs at 
the pNet leaves, and the states are structured as in
the previous section. 
%given by Definition~\ref{def-states}.
To build an open transition one first
 projects the global state into states of the leaves, then applies
pLTS transitions on these states, and compose them with actions of
holes using synchronisation vectors. %The pNet
%structure does not appear in the open-automaton, only the
%set of Holes and the set of Leaves.

%
%\begin{figure}[tb]
%\begin{mathpar}
%  \small
%  \inferrule
%    {\inferrule
%      {
%          \inferrule
%              {1 \xrightarrow {l}_{C_1} 1}
%              {C_1
%                \models
%                \openrule
%                      {1 \xrightarrow{l}_{C_1} 1}
%                      {\ostate{1}\xrightarrow{1}\ostate{1}}
%              }
%        }
%        {\textrm{P}
%          \models
%            \openrule{
%              1 \xrightarrow {l}_{C_1} 1,\,
%              \{\xrightarrow{b_P}_P\},\, b_P\neq l}
%                      {\ostate{1}\xrightarrow{b_P}\ostate{1}}
%        }\\
%%      \sm{fresh}v         \\
%      \inferrule%*[right={L_1}]
%        {
%          \inferrule
%              {0 \xrightarrow{b}_{C_2} 1}
%              {C_4
%                \models
%                \openrule
%                      {0 \xrightarrow{b}_{C_2} 0,\, b\neq r}
%                      {\ostate{0}\xrightarrow{b}\ostate{1}}
%              }
%        }
%        {
%          \textrm{b:Q}\models
%              \openrule
%                  { 0 \xrightarrow {b}_{C_2} 1,\, \ b\neq r}
%                  {\ostate{0}\xrightarrow{b}\ostate{1}}
%        }   
%    }
%    {
%     \textrm{PN1}
%     \models
%        \openrule{
%                       1 \xrightarrow{l}_{C_1} 1 ~~
%                       0 \xrightarrow{b}_{C_2} 1 ~~
%                       \{ \xrightarrow{b_P}_{P} \}               ~~
%                 [ a\neq l \land b\neq r \land b_P=?ch(v) \land b=!ch(v) ]              
%                      }
%     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
%      }\vspace{-4ex}
%\end{mathpar}
%  \caption{Proof of  $OT_{11}$ (with interaction of processes $P$ and action $b$) for ``a.P||b.Q''}
%  \label{usingrules:OT11}
%\end{figure}

\begin{example} \emph{Using the operational rules to compute  open-transitions: }
  in Fig. \ref{usingrules:ot2} we show the deduction tree used to construct and prove the 
  open transition $ot_2$.
\end{example}

\begin{figure}[h]
%  \vspace{-10mm}

\begin{mathpar}
  \small
  \inferrule
    {
          \inferrule
              {0 \xrightarrow {\delta}_{C_1} 0 ~~
               [ s_0=0 ]
              }
              {C_1
                \models
                \openrule
                      {0 \xrightarrow{\delta}_{C_1} 0 ~~
                       [ s_0=0 ] ~~
                       \{ s_0 \leftarrow 1\}
                      }
                      {\ostate{0}\xrightarrow{\delta}\ostate{0}}
              }
	\\
      \inferrule
        {
          \inferrule
              {0 \xrightarrow{acc(x)}_{C_2} 1}
              {C_2
                \models
                \openrule
                      {0 \xrightarrow{acc(x)}_{C_2} 1}
                      {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
              }
        }
        {
          \textrm{acc(x);Q}\models
              \openrule
                  { 0 \xrightarrow {acc(x)}_{C_2} 1}
                  {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
        }   
    }
    {
     \textrm{PN1}
     \models
        \openrule{
                        0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1 ~~
                       \{\xrightarrow{hb_{12}}_P\} ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)] ~~ 
                       \{s_0 \xleftarrow{} 1\}           
                      }
     {\ostate{00} \xrightarrow{\tau} \ostate{01}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of  $ot_{2}$ (interaction of process $P$ and action $acc(x)$) for ``P>>(acc(x);Q)''}
  \label{usingrules:ot2}
\end{figure}

The proof tree uses TR1 twice, for the $\delta$ transition of $C_1$
and for the $acc(x)$ transition of $C_2$, then uses an action
$hb_{12}$ of hole $P$, and combines the results using the first vector
of the PN2 sub-pNet, and the second vector of the top node according
to TR2. This yields a final $\underline{\delta(x)}$ transition. The
deduction tree in the figure shows how the predicates are generated in
this process.

@@

We use a brute-force method to generate all the possible open
transitions in the open automaton, using all possible combinations of
synchronization vectors. Naturally this builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the controller C1 wants to move the control from $P$ to $(acc(x);Q)$,
conducting a $\delta$ transition. However, a synchronization
vector is chosen that does not match with this action, and we can
easily find the contradiction in the 
generated predicate term ``$\OTvar{:ra:11:1}=l \!\OTland\! \delta=\OTvar{:ra:11:1}$''.

@@

@@

>>>>

\item[$\bullet$] $S_j \subseteq \AlgA$ is a set of action terms,
  denoting the $\Sort$\footnote{The formal definition of $Sorts$ (the
    set of actions of a Hole  
  or pNet), and also of $Leaves$ and $Holes$ (respectively all pLTSs
  and all holes in a pNet), can be found in \cite{henrio:Forte2016}.}
  of hole $j$.

@@

The semantics of open pNets will be defined as an open automaton,that
is an automaton where each transition composes transitions 
of several LTSs with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications.

@@

\begin{example}\emph{An open-transition.}
  \TODO{a short explanation}
  \label{OT:lotos-composed}
        \begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
	  \{\xrightarrow{hb}\}, hb=fail(true) \land v=\underline{fail}, \{t := 4\}
        }
	{<s0,t0> \xrightarrow {v} <s1,t1>}
	\end{mathpar}
\end{example}
%
%The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.
%
%The transition in the conclusion of this rule starts in
%state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
%states. It ends in a different structured state, because in $ot_2$
%both pLTS conduct transitions, showing that the control has gone from
%processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
%$acc(x)$. Controller $C_1$ has only one state, but its state variable
%$s_0$ has changed to 1.

\paragraph{Structural Semantic Rules:}
\TODO{I have pushed the rules in the appendix, we need to add a short high
  level desciption here.}

\TODO{ERIC: I choose here to remove the expample of construction of an
  OT as a deduction tree. First because it was already published in
  the Forte'16 paper, but also because it is not the way the algorithm
  is working.}

%% \begin{example} \emph{Using the operational rules to compute  open-transitions: }
%%   in Fig. \ref{usingrules:ot2} we show the deduction tree used to construct and prove the 
%%   open transition $ot_2$.
%% \end{example}

%% \begin{figure}[h]
%% %  \vspace{-10mm}

%% \begin{mathpar}
%%   \small
%%   \inferrule
%%     {
%%           \inferrule
%%               {0 \xrightarrow {\delta}_{C_1} 0 ~~
%%                [ s_0=0 ]
%%               }
%%               {C_1
%%                 \models
%%                 \openrule
%%                       {0 \xrightarrow{\delta}_{C_1} 0 ~~
%%                        [ s_0=0 ] ~~
%%                        \{ s_0 \leftarrow 1\}
%%                       }
%%                       {\ostate{0}\xrightarrow{\delta}\ostate{0}}
%%               }
%% 	\\
%%       \inferrule
%%         {
%%           \inferrule
%%               {0 \xrightarrow{acc(x)}_{C_2} 1}
%%               {C_2
%%                 \models
%%                 \openrule
%%                       {0 \xrightarrow{acc(x)}_{C_2} 1}
%%                       {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
%%               }
%%         }
%%         {
%%           \textrm{acc(x);Q}\models
%%               \openrule
%%                   { 0 \xrightarrow {acc(x)}_{C_2} 1}
%%                   {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
%%         }   
%%     }
%%     {
%%      \textrm{PN1}
%%      \models
%%         \openrule{
%%                         0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1 ~~
%%                        \{\xrightarrow{hb_{12}}_P\} ~~ 
%%                         [s_0=0 \!\land\! hb_{12}=\delta(x)] ~~ 
%%                        \{s_0 \xleftarrow{} 1\}           
%%                       }
%%      {\ostate{00} \xrightarrow{\tau} \ostate{01}}
%%       }\vspace{-4ex}
%% \end{mathpar}
%%   \caption{Proof of  $ot_{2}$ (interaction of process $P$ and action $acc(x)$) for ``P>>(acc(x);Q)''}
%%   \label{usingrules:ot2}
%% \end{figure}

%% The proof tree uses TR1 twice, for the $\delta$ transition of $C_1$
%% and for the $acc(x)$ transition of $C_2$, then uses an action
%% $hb_{12}$ of hole $P$, and combines the results using the first vector
%% of the PN2 sub-pNet, and the second vector of the top node according
%% to TR2. This yields a final $\underline{\delta(x)}$ transition. The
%% deduction tree in the figure shows how the predicates are generated in
%% this process.

@@

We use a brute-force method to generate all the possible open
transitions in the open automaton, using all possible combinations of
synchronization vectors. Naturally this builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
\TODO{Change...} In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the controller C1 wants to move the control from $P$ to $(acc(x);Q)$,
conducting a $\delta$ transition. However, a synchronization
vector is chosen that does not match with this action, and we can
easily find the contradiction in the 
generated predicate term ``$\OTvar{:ra:11:1}=l \!\OTland\! \delta=\OTvar{:ra:11:1}$''.

@@

Before:
  	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
	  \{\xrightarrow{hb}\},\\
\qquad b1=true \land b2=true \land hb=fail(b0) \land b1=b2 \land (b1\cup b2
\implies b0) \land v=\underline{fail}, \{t := 4\}
        }
	{<s0,t0> \xrightarrow {v} <s1,t1>}
	\end{mathpar}

After:
          	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
	  \{\xrightarrow{hb}\}, hb=fail(true) \land v=\underline{fail}, \{t := 4\}
        }
	{<s0,t0> \xrightarrow {v} <s1,t1>}
	\end{mathpar}

@@

\section{pNet semantic rules}

\paragraph{Note:} For convenience, we have included here the semantic rules for pNets,
that were already published in \cite{Forte}.

We build the semantics of an open pNet as an open automaton where
LTSs are the pLTSs at 
the pNet leaves, and the states are structured as in
the previous section. 
%given by Definition~\ref{def-states}.
To build an open transition one first
 projects the global state into states of the leaves, then applies
pLTS transitions on these states, and compose them with actions of
holes using synchronisation vectors. %The pNet
%structure does not appear in the open-automaton, only the
%set of Holes and the set of Leaves.
